### 2020-02-03 defesa/tex/chapter/e3p.tex
wip

Let $E$ represent the coverage area of an ellipse with the given shape parameters $(a, b)$, then a solution of E3PNT can be defined as a pair $(q, \theta) \in \R^2\times[0, \pi)$, such that $\{u, v, w\} \subset \partial E(q, \theta)$. As a last remark, because of its application on \autoref{chapter:ellipses_n}, a method would only be useful for our case if it can encounter every solution of E3PNT. This requirement, of course, makes the problem more challenging.

@@

One of the steps of the method developed in \autoref{chapter:ellipses_n} is to iterate over every solution of E3PNT for every triplet of points. Of course doing that is only possible for a finite number of solutions. Moreover, even if the number of solutions is finite, discovering a bound for it is essential for determining the algorithm's efficiency.

\begin{lema}\label{lema:e3p}
	E3PNT has at most $6$ solutions.
\end{lema}

@@

In this section we show the results of an experiment we made to study the numerical stability of \autoref{algoritmo:e3p}.

We generate random instances of E3PNT according to
\begin{equation}
u, v, w \sim \mathcal{N}(0, K^2),
\end{equation}
with $K\in \R_{>0}$ being a parameter of the experiment. We also define the ellipse's shape parameters ass
\begin{align}
	a = K\\
	b = \frac{K}{2}.
\end{align}
For the experiment, we tried several values for $K$ and adopted the error measure to be $|\xi(\hat{\theta})|$, with $\hat{\theta}$ being a root of the complex polynomial $h$. 
For each different value of $K$, we ran the algorithm for $1000$ random instances. The results are shown in \autoref{fig:e3p_error} where the maximum and the average error  is plotted for each value of $K$ for which the experiment was run.

\begin{figure}[H]
	\centering
	\caption{The interpolation error measured on roots that were found.}
	\includegraphics{tex/figures/e3p_error}
	\fautor
	\label{fig:e3p_error}
\end{figure}

>>>>

Let $E$ represent the coverage area of an ellipse with the given shape parameters $(a, b)$, then a solution of E3PNT can be defined as a pair $(q, \theta) \in \R^2\times[0, \pi)$, such that $\{u, v, w\} \subset \partial E(q, \theta)$. As a last remark, because of its application on \autoref{chapter:mcer}, a method would only be useful for our case if it can encounter every solution of E3PNT. This requirement, of course, makes the problem more challenging.

@@

One of the steps of the method developed in \autoref{chapter:mcer} is to iterate over every solution of E3PNT for every triplet of points. Of course doing that is only possible for a finite number of solutions. Moreover, even if the number of solutions is finite, discovering a bound for it is essential for determining the algorithm's efficiency.

\begin{lema}\label{lema:e3p}
	Any instance of E3PNT has at most $6$ solutions.
\end{lema}

@@

\subsection{Solutions for some instances}

From \autoref{lema:e3p}, we know that any instance of E3P has at most six solutions. At first, though, it seemed like that bound was loose as we could not find any instances that made \autoref{algoritmo:e3p} return six solutions. 
After some investigation, this turned out not to be the case. An example with six solutions is presented in \autoref{fig:e3p_6sols}. It is worth mentioning that the only instances that we found which produce six solutions have their three points form an equilateral triangle.

\begin{figure}[H]
	\centering
	\caption{An instance of E3P with six solutions.}
	\includegraphics[scale=.35]{tex/figures/e3p_6sols}
	\fautor
	\label{fig:e3p_6sols}
\end{figure}

Another example is shown in \autoref{fig:e3p_4sols} where an instance of E3P is displayed along with its four solutions.

\begin{figure}[H]
	\centering
	\caption{An instance of E3P with four solutions.}
	\includegraphics[scale=.35]{tex/figures/e3p_4sols}
	\fautor
	\label{fig:e3p_4sols}
\end{figure}

\subsection{Numerical Stability}

In this section we show the results of an experiment that is developed to study the numerical stability of \autoref{algoritmo:e3p}.

We define $K\in\R_{>0}$ as an experiment parameter and consider instances that the ellipse has shape parameters $(K, \frac{K}{2})$. Then, random instances are created by setting the three points to be $(K\cos(t_j), \frac{K\sin{(t_j)}}{2})$, $j=1..3$, with $t_j$ being uniformly distributed in the interval $[0, 2\pi]$.
To evaluate the stability of \autoref{algoritmo:e3p}, for any solution $(\hat{q}, \hat{\theta})$ found, we adopt its error measure to be $|\xi(\hat{\theta})|$.

The experiment was run for various values of $K$. For each one of them, $100$ randomly generated instances were considered. The results are shown in \autoref{fig:e3p_error}, where the average and the maximum error are plotted.

\begin{figure}[H]
	\centering
	\caption{The error measured on solutions found by \autoref{algoritmo:e3p}.}
	\includegraphics[scale=.9]{tex/figures/e3p_error}
	\fautor
	\label{fig:e3p_error}
\end{figure}

### 2020-02-03 defesa/tex/chapter/mcer.tex
wip

An instance of the non-axis-parallel is defined exactly like the axis-parallel one on \autoref{chapter:ellipses}. It is given by a set of demand points $\Pp=\{p_1, \dots, p_n\}$, $p_j\in\R^2$; a list of weights $\Ww:=\{w_1, \dots, w_n\}$, with $w_j\in\R_{\ge0}$ being the weight of point $p_j$;
and $m$ ellipses given by their shape parameters $\Rr:=\{(a_1, b_1), \dots, (a_m, b_m)\}$, with $(a_j, b_j)\in\R_{>0}^2$ and $a_j>b_j$.
Additionally, to make the text more clear, we define a set of $m$ ellipses as $\E = \{E_1, \dots, E_m\}$, with $E_j\colon \R^2\times\R \mapsto \R^2$ being a function that takes the center and angle of rotation where the $j$-th ellipse is located as input, and returns its coverage region as defined by \autoref{eq:rotated_ellipse_co}.
Lastly, an instance of MCER is defined as a tuple $(\Pp, \Ww, \Rr)$.

@@

	As it was shown on \autoref{chapter:ellipses}, $X$ is a region that is limited by arcs of ellipses. As this region is the non-empty intersection of more than one ellipse, there are at least two of these arcs that encounter at one point, creating a vertex. Selecting any of these vertices as $q_j'$ will make $|\Pp \cap \partial E_j(q_j', \theta_j)| \ge 2$.

@@

	From \autoref{chapter:ellipses_intersection}, we can conclude that $|S_j^{(2)}|\le 2\binom{n}{2} = \bigO(n^2)$.
	By \autoref{lema:e3p}, we have that $|S_j^{(3)}| \le 6\binom{n}{3} = \bigO(n^3)$.
	Therefore, $|S_j|=\bigO(n^3)$, and as $|\Omega(\Pp, \Ww, \Rr)|= |S_1|\times|S_2|\times\dots\times|S_m|$, we have $|\Omega(\Pp, \Ww, \Rr)|=\bigO(n^{3m})$.
\end{proof}

@@

		\Procedure{$MCER_1$}{$\Pp, a, b$}
		\State $S \gets \{\}$

@@

After that, we define \autoref{algoritmo:mcer} which takes an instance $(\Pp, \Ww, \Rr)$ of MCER, and returns an optimal solution for it.
Even though it is based on \autoref{th:mcer}, the set of solutions $\Omega(\Pp, \Ww, \Rr)$ is not explicitly built in \autoref{algoritmo:mcer}. Instead, a complete search is done by backtracking the CLS of every ellipse returned by procedure $MCER_1$.

@@

		\State $ans \gets 0$
		\State $(q_{j}^*\gets 0, \dots, q_m^*\gets0); (\theta_{j}^*\gets 0, \dots, \theta_m^*\gets 0)$ \Comment{Setting to $0$ as a default value.}

		\State $S_j \gets MCER_1(Z, a_j, b_j)$

@@

\begin{corolario}
	\autoref{algoritmo:mcer} takes $\bigO(n^{3m})$ computations and returns an optimal solution.
\end{corolario}

@@

@@

Some improvements can be made in the implementation of \autoref{algoritmo:mcer}.
Instead of calling procedure $MCE_1$ every time in the backtracking routine, it is suggested that it is done outside in a pre-processing phase.

Let $(Q, \Theta)$ and $(Q', \Theta')$ be two solutions of MCER. If, for any $j \in \{1, \dots, m\}$, we have $\Pp \cap E_j(q_j', \theta_j') \subset \Pp \cap E(q_j, \theta_j)$, solution $(Q', \Theta')$ can be dismissed. This can be done in the $MCE_1$ procedure in \autoref{algoritmo:mcer1}.

@@

>>>>

An instance of the non-axis-parallel is defined exactly like the axis-parallel one on \autoref{chapter:mce}. It is given by a set of demand points $\Pp=\{p_1, \dots, p_n\}$, $p_j\in\R^2$; a list of weights $\Ww:=\{w_1, \dots, w_n\}$, with $w_j\in\R_{\ge0}$ being the weight of point $p_j$;
and $m$ ellipses given by their shape parameters $\Rr:=\{(a_1, b_1), \dots, (a_m, b_m)\}$, with $(a_j, b_j)\in\R_{>0}^2$ and $a_j>b_j$.
Additionally, to make the text more clear, we define a set of $m$ ellipses as $\E = \{E_1, \dots, E_m\}$, with $E_j\colon \R^2\times\R \mapsto \R^2$ being a function that takes the center and angle of rotation where the $j$-th ellipse is located as input, and returns its coverage region as defined by \autoref{eq:rotated_ellipse_co}.
Lastly, an instance of MCER is defined as a tuple $(\Pp, \Ww, \Rr)$.

@@

	As it was shown on \autoref{chapter:mce}, $X$ is a region that is limited by arcs of ellipses. As this region is the non-empty intersection of more than one ellipse, there are at least two of these arcs that encounter at one point, creating a vertex. Selecting any of these vertices as $q_j'$ will make $|\Pp \cap \partial E_j(q_j', \theta_j)| \ge 2$.

@@

	From \autoref{section:ellipses_intersection}, we can conclude that $|S_j^{(2)}|\le 2\binom{n}{2} = \bigO(n^2)$.
	By \autoref{lema:e3p}, we have that $|S_j^{(3)}| \le 6\binom{n}{3} = \bigO(n^3)$.
	Therefore, $|S_j|=\bigO(n^3)$, and as $|\Omega(\Pp, \Ww, \Rr)|= |S_1|\times|S_2|\times\dots\times|S_m|$, we have $|\Omega(\Pp, \Ww, \Rr)|=\bigO(n^{3m})$.
\end{proof}

@@

		\Procedure{CLS-MCER}{$\Pp, a, b$}
		\State $S \gets \{\}$

@@

After that, we define \autoref{algoritmo:mcer} which takes an instance $(\Pp, \Ww, \Rr)$ of MCER, and returns an optimal solution for it.
Even though it is based on \autoref{th:mcer}, the set of solutions $\Omega(\Pp, \Ww, \Rr)$ is not explicitly built in \autoref{algoritmo:mcer}. Instead, a complete search is done by backtracking the CLS of every ellipse returned by procedure CLS-MCER.

@@

		\State $ans \gets 0$
		\State $(q_{j}^*, \dots, q_m^*); (\theta_{j}^*, \dots, \theta_m^*) \gets (0, \dots 0); (0, \dots, 0)$ \Comment{Setting to $0$ as a default value.}

		\State $S_j \gets \textnormal{CLS-MCER}(Z, a_j, b_j)$

@@

\begin{corolario}
	\autoref{algoritmo:mcer} takes $\bigO(n^{3m})$ computations and returns an optimal solution for an instance of MCER.
\end{corolario}

@@

In this section, we consider the extended version of MCER where each facility has a cost assigned to it and exactly $k$ of them must be used in a solution. We refer to this version of the problem as \sigla{MCER-$k$}{Maximum Covering by Ellipses with Rotation and a $k$-constraint}.

An instance of MCER-$k$ has the same parameters as MCER plus a list of costs \mbox{$\Cc:=\{c_1, \dots, c_m\}$}, with $c_j\in\R_{>0}$ being the cost of the $j$-th facility; and $k\in\mathbb{N}$, $k\le m$.

A solution for MCER-$k$ is given by $(I, Q, \Theta)$, with $I:=\{i_1, \dots, i_k\}\subset \{1, \dots, m\}$; \\\mbox{$Q:=(q_1, \dots, q_k)\in\R^{2k}$}, with $q_j$ being the center of the $i_j$-th ellipse; and \mbox{$\Theta:=(\theta_1, \dots, \theta_k) \in [0, \pi)^k$}, with $\theta_j$ being the angle of rotation of the $i_j$-th ellipse. An optimal solution of MCER-$k$ is given by the optimization problem

\begin{equation*}
	\max_{I, Q, \Theta} w\left(\bigcup_{j=1}^k \Pp \cap E_{i_j}(q_j, \theta_j)\right).
\end{equation*}

Then, in the same way that it is done in \autoref{chapter:mce}, we introduce \autoref{algoritmo:mcer-k} for MCER-$k$ that uses \autoref{algoritmo:mcer} for every $I:=\{i_1, \dots, i_k\} \subset \{1, \dots, m\}$. Therefore, \autoref{algoritmo:mcer-k} returns an optimal solution for MCER-$k$ in $\bigO(\binom{m}{k}n^{3k}) = \bigO(n^{4m})$ time.

\begin{algoritmo}
	\caption{Algorithm for MCER-$k$}\label{algoritmo:mcer-k}

	\begin{algorithmic}[1]
		\Require{A set of points $\Pp=\{p_1,\dots,p_n\}$, a list of weights $\Ww=\{w_1, \dots, w_n\}$, a list of shape parameters $\Rr=\{(a_1, b_1), \dots, (a_m, b_m)\}$, a list of costs $\Cc=\{c_1, \dots, c_m\}$, and $k\in \mathbb{N}$.}
		\Ensure{An optimal solution for MCER-$k$.}

		\Procedure{MCE-$k$}{$\Pp, \Ww, \Rr, \Cc, k$}
		\State $I^* = \{i_1^*, \dots, i_k^*\}\gets \{1, \dots, k\}$
		\State $Q^* = (q_1^*, \dots, q_k^*) \gets (0, \dots, 0)$
		\State $\Theta^* = (\theta_1^*, \dots, \theta_k^*) \gets (0, \dots, 0)$

		\ForAll{$I=\{i_1, \dots, i_k\} \subset \{1, \dots, m\}$}

		\State $\Rr' \gets \{(a_j, b_j) \in \Rr: j \in I\}$
		\State $(q_1, \dots, q_k); (\theta_1, \dots, \theta_k) \gets MCER(\Pp, \Ww, \Rr')$

		\If{$w(\bigcup_{j=1}^k \Pp \cap E_{i_j}(q_j, \theta_j)) - \sum_{j\in I} c_j > w(\bigcup_{j=1}^k \Pp \cap E_{i_j^*}(q_j^*, \theta_j^*)) - \sum_{j\in I^*}c_{j}$}
		\State $Q^* \gets (q_1, \dots, q_k)$
		\State $\Theta^* \gets (\theta_1, \dots, \theta_k)$
		\State $I^* \gets I$
		\EndIf
		\EndFor

		\State \Return $I^*, Q^*, \Theta^*$
		\EndProcedure
	\end{algorithmic}
\end{algoritmo}

@@

Some improvements can be made in the implementation of \autoref{algoritmo:mcer}.
Instead of calling procedure CLS-MCER every time in the backtracking routine, it is suggested that it is done outside in a pre-processing phase.

Let $(Q, \Theta)$ and $(Q', \Theta')$ be two solutions of MCER. If, for any $j \in \{1, \dots, m\}$, we have $\Pp \cap E_j(q_j', \theta_j') \subset \Pp \cap E(q_j, \theta_j)$, solution $(Q', \Theta')$ can be dismissed. This can be done in the CLS-MCER procedure in \autoref{algoritmo:mcer1}.

@@

I n the next section we present the results of numerical experiments for MCER and MCER-$k$ and compare them with the results published in \citeonline{andreta}.
